---
title: "Using the Chicago Open Data Portal"
author: "Andrew Mehrmann"
date: "November 22, 2016"
output: html_document
---

```{r, echo=F}
knitr::opts_chunk$set(message = F)
```

As with any API, your first step will be setting up an application so you may receive your API token. Head to the [Data Portal](https://data.cityofchicago.org/login) and create an account, then your application.


```{r, message=F}
dataset <- 'jcxq-k9xf'
app_token <- getOption('chidatatoken')
secret_token <- getOption('chidatasecret')


request_url <- sprintf('https://data.cityofchicago.org/resource/%s.csv?$$app_token=%s', dataset, app_token)

payload <- httr::GET(request_url)
df <- httr::content(payload)
head(df)
```

```{r}
library(dplyr)

# now use dplyr for some analysis
request_url %>%
  # get the url
  httr::GET() %>%
  # get the content from the payload
  httr::content() %>% 
  # sort by per capita income
  arrange(desc(per_capita_income_)) %>% 
  # show only the columns we care about
  select(community_area_name, per_capita_income_)
```

The [City of Chicago Github](https://github.com/Chicago) actually includes an R package for making API requests called [RSocrata](https://github.com/Chicago/RSocrata).


Now lets map the Chicago Community areas using [leaflet](https://rstudio.github.io/leaflet/json.html) and a type of file called `geojson`. We can retrieve this from the Chicago Data Portal.

```{r}
library(leaflet)

makeurl <- function(dataset, filetype, token){
  request_url <- sprintf('https://data.cityofchicago.org/resource/%s.%s?$$app_token=%s', dataset, filetype, token)
}

shapefile_url <- makeurl('igwz-8jzy', 'geojson', getOption('chidatatoken'))

payload <- httr::GET(shapefile_url)
geojson <- httr::content(payload, 'text')

leaflet() %>% setView(lng = -87.6298, lat = 41.8781, zoom = 9) %>%
  addTiles() %>%
  addGeoJSON(geojson, weight = 1, color = "#444444", fill = T)
```

```{r}
library(jsonlite)

# From http://data.okfn.org/data/datasets/geo-boundaries-world-110m
geojson_parsed <- geojson %>%
  paste(collapse = "\n") %>%
  fromJSON(simplifyVector = FALSE)

leaflet() %>% setView(lng = -87.6298, lat = 41.8781, zoom = 9) %>%
  addTiles() %>%
  addGeoJSON(geojson_parsed, weight = 1, color = "#444444", fill = T)
```

```{r}
income_df <- makeurl('jcxq-k9xf', 'csv', getOption('chidatatoken')) %>%
  httr::GET() %>%
  # get the content from the payload
  httr::content() %>% 
  arrange(ca)

# get the geojson in list form
geojson_parsed <- makeurl('igwz-8jzy', 'geojson', getOption('chidatatoken')) %>%
  httr::GET() %>%
  httr::content('text') %>%
  paste(collapse = "\n") %>%
  fromJSON(simplifyVector = FALSE) 

# set default stype
geojson_parsed$style <- list(
  weight = 1,
  color = "#555555",
  opacity = 1,
  fillOpacity = 0.5
)

# make a color palette from the income dataframe
pal <- colorQuantile("Greens", income_df$per_capita_income_)

geojson_parsed$features <- 
  geojson_parsed$features %>%
  # set the income feature for each CA
  lapply(function(x) {
    x$properties$income <- as.numeric(df[as.numeric(x$properties$area_num_1), 'per_capita_income_'])
    x
    }) %>%
  # set the fill color based on income and pallete
  lapply(function(feat) {
  feat$properties$style <- list(
    fillColor = pal(
      feat$properties$income
    )
  )
  feat
})


# create the map
leaflet() %>% setView(lng = -87.6298, lat = 41.8781, zoom = 9) %>%
  addTiles() %>%
  addGeoJSON(geojson_parsed)
```



