---
title: "Loops vs. Apply"
author: "Andrew Mehrmann"
date: "September 20, 2016"
output: html_document
---

## Make Some Data

We want to make a long vector of data that we are going to cube (take `x^3` for `x` in `g`). `rnorm` will make both positive and negative floats to give us something to work with.

```{r}
g <- rnorm(100000)
```

## Using Apply

We can apply the function `x^3` to each element in `g` like so:

```{r}
apply_way <- function(g){
  return(sapply(g, function(x) x^3))
}
```

## Looping and Appending the Result

We can also loop through each element in `g` and cube it, then append the result to an empty result vector.

```{r}
loop_way <- function(g){
  res <- c()
  for (i in g){
    res <- c(res, i^3)
  }
  return(res)
}
```


## Looping with Indexing

We can also grab each element in `g` by index and append the result to the result vector at the corresponding index. 

```{r}
loop_way_2 <- function(g){
  num <- length(g)
  res <- c()
  for (i in 1:num){
    res[i] <- g[i]^3
  }
  return(res)
}
```

## Looping with Indexing and Preallocation

Finally, we can *preallocate* memory for the result vector by figuring out how long it is supposed to be before looping.

```{r}
loop_way_3 <- function(g){
  num <- length(g)
  res <- vector(mode="numeric", length=num)
  for (i in 1:num){
    res[i] <- g[i]^3
  }
  return(res)
}
```

## Run the Experiment

```{r, cache=T}
run_sim <- function(g){
  # apply
  res1 <- system.time(ap_res <- apply_way(g))
  # loop-append 
  res2 <- system.time(lo_res <- loop_way(g))
  # loop-index
  res3 <- system.time(lo2_res <- loop_way_2(g))
  # loop-preallocate
  res4 <- system.time(lo3_res <- loop_way_3(g))
  # make a table of results
  tab <- data.frame(apply=res1[3], 
                    append=res2[3], 
                    index=res3[3], 
                    preallocate=res4[3])
  return(tab)
}
tab <- run_sim(g)
```

## Results

<style type="text/css">
.table {

    width: 40%;

}
</style>

```{r,results='asis', echo=F}
knitr::kable(tab)
```

As you can see, preallocation is almost as fast as using `apply` in this case. It is clear why loops in `R` get a bad rap, but most of this is the result of bad code and not an inherent problem with `R`. Always preallocate!

## Bonus - Changing the Length of `g` to 10000

```{r, results='asis'}
h <- rnorm(10000)
tab2 <- run_sim(h)
knitr::kable(tab2)
```

The results are *roughly* linear in this case. A vector 10% as long takes a little less than 10% of the time to cube for each method.
